---
title: "Global Disease Burden Analysis"
author: "Giulia Hofer" # AUTHOR NAMES HERE!
output: github_document
---

```{r include = F}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

This document compares the Disability-Adjusted Life Year (DALY) diseases burden for three countries: United States, Brazil, and India. We compare the burden due to communicable, maternal, neonatal, and nutritional diseases (CMNN), non-communicable diseases (NCDs), and overall disease burden for these countries. 

The source data is from the Institute for Health Metrics and Evaluation (IHME) Global Burden of Disease (GBD) study. 

```{r child='cmnn_burden.Rmd'}
```

```{r child='ncd_burden.Rmd'}
```

```{r child='overall_burden.Rmd'}
```


## Plot of cumulative DALY burden for the three countries over time, for each of the three burden estimate types

```{r}
# Load necessary libraries
library(dplyr)

data_cmnn_countries$burden_type <- 'Communicable'
data_ncd_3_countries$burden_type <- 'Noncommunicable'
data_overall_countries$burden_type <- 'All-causes'

data_combined <- bind_rows(data_cmnn_countries, data_ncd_3_countries, data_overall_countries)

# Calculate cumulative DALY for each country and burden type
data_cumulative <- data_combined %>%
  group_by(Entity, burden_type) %>%
  arrange(Entity, burden_type, Year) %>%
  mutate(cumulative_DALY = cumsum(DALYs))

# Plotting cumulative DALY burden
ggplot(data = data_cumulative, aes(x = Year, y = cumulative_DALY, color = Entity, group = interaction(Entity, burden_type))) + 
  geom_line() +
  facet_wrap(~ burden_type, scales = "free_y") +
  labs(title = "Cumulative DALY Burden Over Time",
       x = "Year",
       y = "Cumulative DALY Burden",
       color = "Country") +
  theme_minimal()

```
