# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data", "burden-of-disease-cmnn.csv"), show_col_types = FALSE)
```

## Table of Estimates for CMNN Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.

data_cmnn_countries <- data_cmnn %>% filter(Entity %in% c("United States", "India", "Brazil")) %>% rename("DALYs" = "DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)") %>% select(Entity, Year, DALYs)

# Pivot data to wide format
data_wide <- data_cmnn_countries %>%
  pivot_wider(names_from = Entity, values_from = DALYs) # Ensure 'country' and 'DALY' match your column names

# Render the table with kable
kable(data_wide)
```

## Plot Showing Trends in CMNN Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

# Load the necessary library
library(ggplot2)

# Plotting the DALY burden over time

ggplot(data = data_cmnn_countries, aes(x = Year, y = DALYs, color = Entity)) + 
  geom_line() +
  labs(title = "Trend of DALY Burden Over Time in communicable diseases",
       x = "Year",
       y = "DALY Burden",
       color = "Country") +
  theme_minimal()


```

## Summary of CMNN Burden Findings

The chart shows a consistent decline in the DALY burden from communicable diseases over time in India, indicating substantial improvements in managing these diseases. Brazil's DALY burden is also decreasing, albeit at a slower rate than India's. Meanwhile, the United States displays a relatively flat trend, suggesting a stable and low DALY burden from communicable diseases throughout the period. This may reflect the impact of sustained public health efforts and healthcare advancements in these regions.


