# Overall Disease Burden

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the overall disease data
data_overall <- read_csv(here("data", "burden-of-disease-all-causes.csv"))
```

## Table of Estimates for Overall Disease Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.


data_overall_countries <- data_overall %>% filter(Entity %in% c("United States", "India", "Brazil")) %>% rename("DALYs" = "DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)") %>% select(Entity, Year, DALYs)

# Pivot data to wide format
data_wide <- data_overall_countries %>%
  pivot_wider(names_from = Entity, values_from = DALYs) # Ensure 'country' and 'DALY' match your column names

# Render the table with kable
kable(data_wide)

```

## Plot Showing Trends in Overall Disease Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

# Load the necessary library
library(ggplot2)

# Plotting the DALY burden over time

ggplot(data = data_overall_countries, aes(x = Year, y = DALYs, color = Entity)) + 
  geom_line() +
  labs(title = "Trend of all-cause DALY Burden Over Time",
       x = "Year",
       y = "DALY Burden",
       color = "Country") +
  theme_minimal()
```

## Summary of Overall Disease Burden Findings

The chart indicates a significant decrease in the all-cause DALY burden over time for India, showing a strong positive trend in overall health or effectiveness of health interventions. Brazil also shows a declining trend, but with a less steep slope than India, suggesting slower progress. The United States exhibits the lowest DALY burden of the three and maintains a relatively stable level over the years, possibly reflecting an established healthcare system effectively managing the overall health burden.

